Cell Decompositions for Rank Two Quiver Grassmannians

### 2017-11-20 Grassmannian_cells.tex
Removed some statements about GL_n action

\begin{lemma}
  \footnote{Is this true?}
  The reflection functors $\Sigma_i:\rep K(n)\to\rep K(n)$ are $GL_n(\CC)$-equivariant. 
\end{lemma}

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.
The following is well-known.
\begin{theorem}
  \label{th:rigids}
  For each $m\ge1$, there exist unique (up to isomorphism) indecomposable rigid representations $P_m$ and $I_m$ of $K(n)$ with dimension vectors $(u_m,u_{m-1})$ and $(u_{m-1},u_m)$ respectively. 
  Moreover, any rigid representation of $K(n)$ is isomorphic to one of the form $P_m^{a_1}\oplus P_{m+1}^{a_2}$ or $I_m^{a_1}\oplus I_{m+1}^{a_2}$ for some $m\ge1$ and some $a_1,a_2\ge0$.
\end{theorem}
The representations $P_m$ are called the \emph{preprojective} representations of $K(n)$ and the representations $I_m$ are called \emph{preinjective}.
\begin{remark}
  \label{rem:reflection recursion}
  These representations satisfy the following recursions using the reflection functors:
  \[P_1=S_1,\quad P_m=\Sigma_2 P_{m-1},\quad I_1=S_2,\quad I_m=\Sigma_1 I_{m-1}\]
  for $m\ge2$.
  In particular, we have $P_{m-1}=\tau P_{m+1}$ and $I_{m+1}=\tau I_{m-1}$ for $m\ge2$. 
\end{remark}
\begin{corollary}
  \label{cor:preprojective stability}
  For $m\ge1$ and $g\in G$, the representations $P_m$ and $g.P_m$ are isomorphic.
\end{corollary}
\begin{proof}
  By construction, the representation $g.P_m$ also has dimension vector $(u_m,u_{m-1})$ and is indecomposable by Corollary~\ref{cor:indecomposability}.
  Thus by Theorem~\ref{th:rigids}, $P_m$ and $g.P_m$ are isomorphic.
\end{proof}
For $g\in GL_n(\CC)$, write $\phi_{m,g}:P_m\to g.P_m$ for the (unique up to scalars) isomorphism from Corollary~\ref{cor:preprojective stability}.

Set $H_m:=\Hom(P_m,P_{m+1})$ for $m\ge1$ and define an action of $GL_n(\CC)$ on $H_m$ by $g.\theta:=\phi_{m+1,g}^{-1}\circ\theta^g\circ\phi_{m,g}$.
Note that the first $\theta$ here represents an element of $H_m$ while the second $\theta^g$ represents the corresponding element of $H_m^g:=\Hom(g.P_m,g.P_{m+1})$ from Lemma~\ref{le:hom equivariance}.

Write $Gr(H_m)$ for the \emph{total Grassmannian} of $H_m$ whose elements are nontrivial proper subspaces $V\subset H_m$.
Some results below remain true if we allow $H_m$ or $0$ as elements of $Gr(H_m)$, but not all, so for uniformity of exposition we omit these possibilities.
Lemma~\ref{le:hom equivariance} gives an identification of $Gr(H_m)$ with $Gr(H_m^g)$ and we write $V\mapsto V^g$ for this correspondence.
For $1\le d\le n-1$, let $Gr_d(H_m)\subset Gr(H_m)$ be the subvariety of $d$-dimensional subspaces of $H_m$.
\begin{lemma}
  \label{le:transitivity}
  The action of $GL_n(\CC)$ on $H_m$ induces a transitive action of $GL_n(\CC)$ on each component $Gr_d(H_m)$.
\end{lemma} 
\begin{proof}

\end{proof}
\begin{lemma}
  \label{le:injective evaluation maps}
  For any $V\in Gr(H_m)$, the natural evaluation map $ev_V:P_m\otimes V\to P_{m+1}$ is injective.
\end{lemma}
%%%%%%%%%%%%%%%%%%%
% needs polishing %
%%%%%%%%%%%%%%%%%%%
\begin{proof}
  \footnote{This needs to be cleaned up, the proof seems false as written.}
  A nonzero element $f\in V$ corresponds to an irreducible morphism $f:P_m\to P_{m+1}$ and thus does not factor through any representation $Z\ne P_m$, in particular $f$ is injective.
  But any nonzero element of $\ker ev_V$ gives rise to a nonzero $f\in V$ which is not injective and thus $ev_V$ must be an injective map.
\end{proof}

In what follows we will not distinguish between $P_m\otimes V$ and its image under $ev_V$.
Similarly, for $V^g\in Gr(H_m^g)$, we identify $g.P_m\otimes V^g$ with its image in $g.P_{m+1}$ under the natural evaluation map. 
\begin{lemma}
  \label{le:g equalities}
  For any $g\in GL_n(\CC)$, the subrepresentation $g.(P_m\otimes V)\subset g.P_{m+1}$ is equal to $g.P_m\otimes V^g\subset g.P_{m+1}$ and the subrepresentation $\phi_{m+1,g}^{-1}(g.P_m\otimes V^g)\subset P_{m+1}$ is equal to $P_m\otimes g.V\subset P_{m+1}$.
\end{lemma}
\begin{proof}
  The first equality is immediate from the definitions.
  The second equality also follows from the definitions by observing that $\phi_{m+1,g}(P_m\otimes g.V)=g.P_m\otimes V^g$.
\end{proof}

Following Lemma~\ref{le:injective evaluation maps} we define the following.
\begin{definition}
  For $V\in Gr(H_m)$, define the \emph{truncated preprojective} $P_{m+1}^V$ to be the cokernel of the map $ev_V:P_m\otimes V\to P_{m+1}$, i.e.\ we have a short exact sequence
  \begin{equation}
    \label{eq:truncated preprojectives}
    0\longrightarrow P_m\otimes V\stackrel{ev_V}{\longrightarrow} P_{m+1}\stackrel{\pi_V}{\longrightarrow} P_{m+1}^V\longrightarrow 0.
  \end{equation}
\end{definition}
\begin{remark}
  It will be convenient to also set $P_{m+1}^0=P_{m+1}$, observe that this notation is consistent with taking $V=0$ in the sequence \eqref{eq:truncated preprojectives}.
\end{remark}

Any subrepresentation $E\subset P_{m+1}^V$ corresponds by taking pullbacks to a chain of inclusions $P_m\otimes V\subset\tilde E\subset P_{m+1}$.
\begin{theorem}
  The quiver Grassmannian $Gr_\bfe(P_{m+1}^V)$ is smooth for each dimension vector $\bfe$ and each $V\in Gr(H_m)$.
  Moreover, for $V,W\in Gr_d(H_m)$, we have $Gr_\bfe(P_{m+1}^V)\cong Gr_\bfe(P_{m+1}^W)$ for all dimension vectors $\bfe$.
\end{theorem}
\begin{proof}
  Both claims are proven via the same geometric construction.
  For each dimension vector $\bfe$, there is a smooth variety
  \[X_{\bfe,m}=\{(\tilde E,V):V\in Gr_d(H_m),\ P_m\otimes V\subset\tilde E\subset P_{m+1}i,\ \udim\tilde E=\bfe+d\udim P_m\}.\]
  Under the natural projection $\pi:X_{\bfe,m}\to Gr_d(H_m)$, the fiber $\pi^{-1}(V)$ identifies with $Gr_\bfe(P_{m+1}^V)$.
  In particular, this implies $Gr_\bfe(P_{m+1}^V)$ is smooth for each $V\in Gr_d(H_m)$.

  The variety $X_{\bfe,m}$ supports an action of $GL_n(\CC)$ by setting $g.(\tilde E,V)=\big(\phi_{m+1,g}^{-1}(g.\tilde E),g.V\big)$.
  Indeed, by Lemma~\ref{le:g equalities} we have 
  \[P_m\otimes g.V=\phi_{m+1,g}^{-1}(g.P_m\otimes V^g)=\phi_{m+1,g}^{-1}\big(g.(P_m\otimes V)\big)\subset\phi_{m+1,g}^{-1}(g.\tilde E).\]
  Observe that the projection map $\pi$ is $GL_n(\CC)$-equivariant and thus by Lemma~\ref{le:transitivity} the quiver Grassmannians $Gr_\bfe(P_{m+1}^V)$ and $Gr_\bfe(P_{m+1}^W)$ are isomorphic for any $V,W\in Gr_d(H_m)$.
\end{proof}

>>>>

While the reflection functors are not $GL_n(\CC)$-equivariant, they do admit the following twisted equivariance.
\begin{lemma}
  For $g\in G$, the reflection functors $\Sigma_i:\rep K(n)\to\rep K(n)$ satisfy $g.\Sigma_i(M)=\Sigma_i(g^{-T}.M)$. 
\end{lemma}
\begin{proof}
  We present all the details for $\Sigma_2$, the proof for $\Sigma_1$ is similar.
  For $M\in\rep K(n)$, we have $\Sigma_2(M)=(M_1,M'_2,M'_{a_i})$, where $M'_2$ fits into the following exact sequence:
  \[\xymatrix{0 \ar[r] & M_2 \ar[r]^{\bigoplus\limits_{i=1}^n M_{a_i}} & \bigoplus\limits_{i=1}^n M_1 \ar[r]^\pi & M'_2 \ar[r] & 0}\]
  and $M'_{a_i}=\pi\circ\iota_i$ for $\iota_i:M_1\to\bigoplus_{i=1}^n M_1$ the inclusion of the $i$-th factor.
  Then for $g=(g_{ij})\in GL_n(\CC)$, we have $g.\Sigma_2(M)=(M_1,M'_2,M''_{a_i})$ with 
  \[M''_{a_i}=\sum\limits_{j=1}^n g_{ij}M'_{a_j}=\sum\limits_{j=1}^n g_{ij}(\pi\circ\iota_j)=\pi\circ\sum\limits_{j=1}^n g_{ij}\iota_j=\pi\circ g^T\circ\iota_i.\]
  In particular, we may construct $g.\Sigma_2(M)$ using the following exact sequence:
  \[\xymatrix{0 \ar[r] & M_2 \ar[r]^{g^{-T}\circ\bigoplus\limits_{i=1}^n M_{a_i}} & \bigoplus\limits_{i=1}^n M_1 \ar[r]^{\pi\circ g^T} & M'_2 \ar[r] & 0},\]
  i.e. $g.\Sigma_2(M)=\Sigma_2(g^{-T}.M)$.
\end{proof}

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.
The following is well-known.
\begin{theorem}
  \label{th:rigids}
  For each $m\ge1$, there exist unique (up to isomorphism) indecomposable rigid representations $P_m$ and $I_m$ of $K(n)$ with dimension vectors $(u_m,u_{m-1})$ and $(u_{m-1},u_m)$ respectively. 
  Moreover, any rigid representation of $K(n)$ is isomorphic to one of the form $P_m^{a_1}\oplus P_{m+1}^{a_2}$ or $I_m^{a_1}\oplus I_{m+1}^{a_2}$ for some $m\ge1$ and some $a_1,a_2\ge0$.
\end{theorem}
The representations $P_m$ are called the \emph{preprojective} representations of $K(n)$ and the representations $I_m$ are called \emph{preinjective}.
\begin{remark}
  \label{rem:reflection recursion}
  These representations satisfy the following recursions using the reflection functors:
  \[P_1=S_1,\quad P_m=\Sigma_2 P_{m-1},\quad I_1=S_2,\quad I_m=\Sigma_1 I_{m-1}\]
  for $m\ge2$.
  In particular, we have $P_{m-1}=\tau P_{m+1}$ and $I_{m+1}=\tau I_{m-1}$ for $m\ge2$. 
\end{remark}
%\begin{corollary}
%  \label{cor:preprojective stability}
%  For $m\ge1$ and $g\in G$, the representations $P_m$ and $g.P_m$ are isomorphic.
%\end{corollary}
%\begin{proof}
%  By construction, the representation $g.P_m$ also has dimension vector $(u_m,u_{m-1})$ and is indecomposable by Corollary~\ref{cor:indecomposability}.
%  Thus by Theorem~\ref{th:rigids}, $P_m$ and $g.P_m$ are isomorphic.
%\end{proof}
%For $g\in GL_n(\CC)$, write $\phi_{m,g}:P_m\to g.P_m$ for the (unique up to scalars) isomorphism from Corollary~\ref{cor:preprojective stability}.

Set $H_m:=\Hom(P_m,P_{m+1})$ for $m\ge1$. %and define an action of $GL_n(\CC)$ on $H_m$ by $g.\theta:=\phi_{m+1,g}^{-1}\circ\theta^g\circ\phi_{m,g}$.
%Note that the first $\theta$ here represents an element of $H_m$ while the second $\theta^g$ represents the corresponding element of $H_m^g:=\Hom(g.P_m,g.P_{m+1})$ from Lemma~\ref{le:hom equivariance}.
Write $Gr(H_m)$ for the \emph{total Grassmannian} of $H_m$ whose elements are nontrivial proper subspaces $V\subset H_m$.
Some results below remain true if we allow $H_m$ or $0$ as elements of $Gr(H_m)$, but not all, so for uniformity of exposition we omit these possibilities.
%Lemma~\ref{le:hom equivariance} gives an identification of $Gr(H_m)$ with $Gr(H_m^g)$ and we write $V\mapsto V^g$ for this correspondence.
For $1\le d\le n-1$, let $Gr_d(H_m)\subset Gr(H_m)$ be the subvariety of $d$-dimensional subspaces of $H_m$.
%\begin{lemma}
%  \label{le:transitivity}
%  The action of $GL_n(\CC)$ on $H_m$ induces a transitive action of $GL_n(\CC)$ on each component $Gr_d(H_m)$.
%\end{lemma} 
%\begin{proof}
%  The map $\theta\mapsto g.\theta$ for $\theta\in H_m$ induces a group homomorphism $GL_n(\CC)\to GL(H_m)$ which we denote by $g\mapsto\tilde g_m$.
%  We claim that this homomorphism is surjective which immediately implies the result.
%
%  We work by induction on $m$.
%  Suppose $GL_n(\CC)\to GL(H_m)$ is surjective.
%
%\end{proof}
\begin{lemma}
  \label{le:injective evaluation maps}
  For any $V\in Gr(H_m)$, the natural evaluation map $ev_V:P_m\otimes V\to P_{m+1}$ is injective.
\end{lemma}
%%%%%%%%%%%%%%%%%%%
% needs polishing %
%%%%%%%%%%%%%%%%%%%
\begin{proof}
  \footnote{This needs to be cleaned up, the proof seems false as written.}
  A nonzero element $f\in V$ corresponds to an irreducible morphism $f:P_m\to P_{m+1}$ and thus does not factor through any representation $Z\ne P_m$, in particular $f$ is injective.
  But any nonzero element of $\ker ev_V$ gives rise to a nonzero $f\in V$ which is not injective and thus $ev_V$ must be an injective map.
\end{proof}

In what follows we will not distinguish between $P_m\otimes V$ and its image under $ev_V$.
%Similarly, for $V^g\in Gr(H_m^g)$, we identify $g.P_m\otimes V^g$ with its image in $g.P_{m+1}$ under the natural evaluation map. 
%\begin{lemma}
%  \label{le:g equalities}
%  For any $g\in GL_n(\CC)$, the subrepresentation $g.(P_m\otimes V)\subset g.P_{m+1}$ is equal to $g.P_m\otimes V^g\subset g.P_{m+1}$ and the subrepresentation $\phi_{m+1,g}^{-1}(g.P_m\otimes V^g)\subset P_{m+1}$ is equal to $P_m\otimes g.V\subset P_{m+1}$.
%\end{lemma}
%\begin{proof}
%  The first equality is immediate from the definitions.
%  The second equality also follows from the definitions by observing that $\phi_{m+1,g}(P_m\otimes g.V)=g.P_m\otimes V^g$.
%\end{proof}
Following Lemma~\ref{le:injective evaluation maps} we define the following.
\begin{definition}
  For $V\in Gr(H_m)$, define the \emph{truncated preprojective} $P_{m+1}^V$ to be the cokernel of the map $ev_V:P_m\otimes V\to P_{m+1}$, i.e.\ we have a short exact sequence
  \begin{equation}
    \label{eq:truncated preprojectives}
    0\longrightarrow P_m\otimes V\stackrel{ev_V}{\longrightarrow} P_{m+1}\stackrel{\pi_V}{\longrightarrow} P_{m+1}^V\longrightarrow 0.
  \end{equation}
\end{definition}
\begin{remark}
  It will be convenient to also set $P_{m+1}^0=P_{m+1}$, observe that this notation is consistent with taking $V=0$ in the sequence \eqref{eq:truncated preprojectives}.
\end{remark}

%Any subrepresentation $E\subset P_{m+1}^V$ corresponds by taking pullbacks to a chain of inclusions $P_m\otimes V\subset\tilde E\subset P_{m+1}$.
\begin{theorem}
  The quiver Grassmannian $Gr_\bfe(P_{m+1}^V)$ is smooth for each dimension vector $\bfe$ and each $V\in Gr(H_m)$.
  Moreover, for $V,W\in Gr_d(H_m)$, we have $Gr_\bfe(P_{m+1}^V)\cong Gr_\bfe(P_{m+1}^W)$ for all dimension vectors $\bfe$.
\end{theorem}
\begin{proof}
  \footnote{The proof using moduli spaces needs to be written}
  %Both claims are proven via the same geometric construction.
  %For each dimension vector $\bfe$, there is a smooth variety
  %\[X_{\bfe,m}=\{(\tilde E,V):V\in Gr_d(H_m),\ P_m\otimes V\subset\tilde E\subset P_{m+1}i,\ \udim\tilde E=\bfe+d\udim P_m\}.\]
  %Under the natural projection $\pi:X_{\bfe,m}\to Gr_d(H_m)$, the fiber $\pi^{-1}(V)$ identifies with $Gr_\bfe(P_{m+1}^V)$.
  %In particular, this implies $Gr_\bfe(P_{m+1}^V)$ is smooth for each $V\in Gr_d(H_m)$.

  %The variety $X_{\bfe,m}$ supports an action of $GL_n(\CC)$ by setting $g.(\tilde E,V)=\big(\phi_{m+1,g}^{-1}(g.\tilde E),g.V\big)$.
  %Indeed, by Lemma~\ref{le:g equalities} we have 
  %\[P_m\otimes g.V=\phi_{m+1,g}^{-1}(g.P_m\otimes V^g)=\phi_{m+1,g}^{-1}\big(g.(P_m\otimes V)\big)\subset\phi_{m+1,g}^{-1}(g.\tilde E).\]
  %Observe that the projection map $\pi$ is $GL_n(\CC)$-equivariant and thus by Lemma~\ref{le:transitivity} the quiver Grassmannians $Gr_\bfe(P_{m+1}^V)$ and $Gr_\bfe(P_{m+1}^W)$ are isomorphic for any $V,W\in Gr_d(H_m)$.
\end{proof}
